---
title: "CapstoneMarkdown"
output:
  word_document: default
  html_document: default
---

Reading data in R
```{r reading data in R}

library(readxl)

explore <- read_xlsx("Capstone Copy.xlsx")

```

col names
```{r col names}

# colnames(explore)
# 
# unique(explore$Race)
# 
# unique(explore$RaceSub)
# 
# unique(explore$`PHQ2 Dx`)
# 
# class(explore$AgeAsOf12312018)
# 
# unique(explore$AgeAsOf12312018)

```


Filtered
```{r filtered}

library(tidyverse)

explore %>%
    group_by(`PHQ2 Dx`, `PHQ9 Dx`) %>%
    summarize(count = n()) %>%
    spread(`PHQ9 Dx`, count)

##Total patients 6365
##Race removed 764 
##PHQ2/9 removed 

datafiltered <- explore %>%
  filter(Race == "Asian" | Race == "Asian|Unspecified" | Race == "Chinese",
         RaceSub == "Chinese" | RaceSub == "Chinese|Undetermined" | RaceSub == "Chinese|Other", 
         !is.na(`PHQ2 Score`),
         !(`PHQ2 Dx`== "Positive" & is.na(`PHQ9 Dx`)),
         HasADO == 0)

# phq29filter <- explore %>%
#   filter(!is.na(`PHQ2 Score`),
#          !(`PHQ2 Dx`== "Positive" & is.na(`PHQ9 Dx`)),
#          HasADO == 0)

# Racefilter <- explore %>%
#   filter(Race == "Asian" | Race == "Asian|Unspecified" | Race == "Chinese",
#   RaceSub == "Chinese" | RaceSub == "Chinese|Undetermined" | RaceSub == "Chinese|Other")

##Code for filtering PHQ2 Negatives that became PHQ9 Positives and Negatives datafiltered <- datafiltered %>% filter(!(`PHQ2 Dx` == "Negative" & (`PHQ9 Dx` == "Positive" | `PHQ9` Dx == "Negtive")))


PHQ2Neg <- datafiltered %>%
  filter(`PHQ2 Dx` == "Negative")

PHQ9NA <- datafiltered %>%
  filter(is.na(`PHQ9 Dx`))

##code for changing all PHQ9 NA values to Negative -- previously filtered all PHQ2 Negatives

datafiltered$`PHQ9 Dx` = ifelse(is.na(datafiltered$`PHQ9 Dx`), 
                                "Negative", datafiltered$`PHQ9 Dx`)

unique(datafiltered$`PHQ9 Dx`)

table(datafiltered$`PHQ9 Dx`)

```
table function
```{r table function}

table1fx <- function(data, col) {
  byoutcome <- data %>%
    group_by(!!col, `PHQ9 Dx`) %>%
    summarize(count = n()) %>%
    spread(`PHQ9 Dx`, count)
  

byoutcome$Total = apply(byoutcome[2:ncol(byoutcome)], 1, sum, na.rm = TRUE)

byoutcome$ColPercentage = byoutcome$Total/sum(byoutcome$Total,na.rm = TRUE)

byoutcome$PercentPos =
  (byoutcome$Positive/(byoutcome$Positive+byoutcome$Negative))*100

byoutcome$PercentNegative =
  (byoutcome$Negative/(byoutcome$Positive+byoutcome$Negative))*100

byoutcome <- byoutcome[, c(1, 4, 5, 3, 6, 2, 7)]

  return(byoutcome)
}

# table1fx <- function(data, col) {
#   byoutcome <- data %>% 
#     group_by(!!col, `PHQ9 Dx`) %>% 
#     summarize(count = n()) %>% 
#     spread(`PHQ9 Dx`, sum)
#   
#   byoutcome <- byoutcome[, c(1, 3, 2)]
# 
# byoutcome$ColumnPercentage =
#   (byoutcome$PercentPositive)+(byoutcome$PercentNegative)/(by)
#    
# byoutcome$PercentNegative =
#   (byoutcome$Negative/(byoutcome$Positive+byoutcome$Negative))*100
# 
# byoutcome$PercentPositive =
#   (byoutcome$Positive/(byoutcome$Positive+byoutcome$Negative))*100
# 
#   return(byoutcome)
# }


```
Age 
```{r age}

agetable1 <-table1fx(datafiltered, quo(`AgeAsOf12312018`))

agetable1


```
sex 
```{r sex}

sextable1 <- table1fx(datafiltered, quo(`BirthSex`))

sextable1


```

Gender Identity
```{r Gender Identity}

gendertable1 <- table1fx(datafiltered, quo(`Gender Identity`))

gendertable1


```
Sexual Orientation
```{r Sexual Orientation}

sexualorientationtable1 <- table1fx(datafiltered, quo(`Sexual Orientation`))

sexualorientationtable1

```

Birth Country
```{r Birth Country}

datafiltered2 <- datafiltered

datafiltered2$`Birth Country` = ifelse(!datafiltered2$`Birth Country` == "USA",
                                "Foreign Born", datafiltered2$`Birth Country`)

birthcountrytable1 <- table1fx(datafiltered2, quo(`Birth Country`))

birthcountrytable1

```
Age of immigration 
```{r Age of immigration}

datafiltered3 <- datafiltered2

Yearentered <- select(datafiltered2, StudyID, `Year Entered USA`, AgeAsOf12312018)

datafiltered2$`Year Entered USA`[datafiltered2$StudyID == 4815] = 2007
datafiltered2$`Year Entered USA`[datafiltered2$StudyID == 1186] = 2006
datafiltered2$`Year Entered USA`[datafiltered2$StudyID == 4815] = 2005

##turns text into NA, otherwise make character into numeric
datafiltered2$`Year Entered USA`= ifelse(is.na(as.numeric(datafiltered2$`Year Entered USA`)) == TRUE, NA, as.numeric(datafiltered2$`Year Entered USA`))

datafiltered2$`Year Entered USA`= ifelse(datafiltered2$`Year Entered USA`>2018, NA, datafiltered2$`Year Entered USA`)


immigrateyeartable1 <- table1fx(datafiltered2, quo(`Year Entered USA`))

immigrateyeartable1


### Age - (2018 - Year immigrated), coerce new column with the dollar symbol 

datafiltered2$ageofimmigration <- datafiltered2$AgeAsOf12312018 - (2018 - datafiltered2$`Year Entered USA`)

datafiltered2$ageofimmigration = ifelse(datafiltered2$ageofimmigration <=0 , NA, datafiltered2$ageofimmigration)


ageimmigtable1 <- table1fx(datafiltered2, quo(`ageofimmigration`))

ageimmigtable1

langtable1 <- table1fx(datafiltered2, quo(`Preferred Language`))

langtable1

```

Marijuana use
```{r Marijuana use}

marijuanatable1 <- table1fx(datafiltered2, quo(`Marijuana Use`))

marijuanatable1


```
Alcohol Use
```{r Alcohol Use}


datafiltered2$`Alcohol Use` = ifelse(datafiltered2$`Alcohol Use` == "Former", "No", datafiltered2$`Alcohol Use`)

datafiltered2$`Alcohol Use` = ifelse(datafiltered2$`Alcohol Use` == "no", "No", datafiltered2$`Alcohol Use`)

datafiltered2$`Alcohol Use` = ifelse(datafiltered2$`Alcohol Use` == "Negative", "No", datafiltered2$`Alcohol Use`)  

datafiltered2$`Alcohol Use` = ifelse(datafiltered2$`Alcohol Use` == "Never", "No", datafiltered2$`Alcohol Use`)

datafiltered2$`Alcohol Use` = ifelse(datafiltered2$`Alcohol Use` == "yes", "Yes", datafiltered2$`Alcohol Use`)

alcoholtable1 <- table1fx(datafiltered2, quo(`Alcohol Use`))

alcoholtable1

```
Alcohol & Marijuana

```{r Alcohol & Marijuana}

datafiltered2$alcmarj = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`Marijuana Use` == "No", "No Alc or No Marj", NA)

datafiltered2$alcmarj = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`Marijuana Use` == "No", "Yes Alc No Marj", datafiltered2$alcmarj)

datafiltered2$alcmarj = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`Marijuana Use` == "Yes", "No Alc Yes Marj", datafiltered2$alcmarj)

datafiltered2$alcmarj = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`Marijuana Use` == "Yes", "Yes Alc Yes Marj", datafiltered2$alcmarj)

unique(datafiltered2$alcmarj)

alcmarjtable1 <- table1fx(datafiltered2, quo(alcmarj))

alcmarjtable1


```

Other drugs 
```{r Other drugs}

datafiltered2$`Drug Use` = ifelse(datafiltered2$`Drug Use` == "yes", "Yes", datafiltered2$`Drug Use`)

datafiltered2$`Drug Use` = ifelse(datafiltered2$`Drug Use` == "no", "No", datafiltered2$`Drug Use`)

drugusetable1 <- table1fx(datafiltered2, quo(`Drug Use`))
drugusetable1

```

Alcohol & Other Drugs

```{r Alcohol & Other Drugs}

datafiltered2$alcdrugs = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`Drug Use` == "No", "No Alc or No Drug Use", NA)

datafiltered2$alcdrugs = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`Drug Use` == "No", "Yes Alc No Drug Use", datafiltered2$alcdrugs)

datafiltered2$alcdrugs = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`Drug Use` == "Yes", "No Alc Yes Drug Use", datafiltered2$alcdrugs)

datafiltered2$alcdrugs = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`Drug Use` == "Yes", "Yes Alc Yes Drug Use", datafiltered2$alcdrugs)

unique(datafiltered2$alcdrugs)

alcdrugstable1 <- table1fx(datafiltered2, quo(alcdrugs))

alcdrugstable1


```


E-cigarettes
```{r E-cigarettes}

datafiltered2$Ecigarette = ifelse(datafiltered2$Ecigarette == "E-cigarettes", "Yes", datafiltered2$Ecigarette)





ecigtable1 <- table1fx(datafiltered2, quo(Ecigarette))

ecigtable1

```
Smoking Status
```{r Smoking Status}


datafiltered2$`Smoking Status` = case_when(
  datafiltered2$`Smoking Status` == "Current every day smoker" ~ "Current Smoker",
  datafiltered2$`Smoking Status` == "Current some day smoker" ~ "Current Smoker",
  datafiltered2$`Smoking Status` == "Light tobacco smoker" ~ "Current Smoker",
  datafiltered2$`Smoking Status` == "Former smoker" ~ "Former Smoker",
  datafiltered2$`Smoking Status` == "Never smoker" ~ "Never Smoker",
  datafiltered2$`Smoking Status` == "No" ~ "Never Smoker")




smoketable1 <- table1fx(datafiltered2, quo(`Smoking Status`))

smoketable1 <- smoketable1[c(3,2,1,4),]

smoketable1

```



CRAFFT Score
```{r Crafft Score}

datafiltered2$`CRAFFT Score` = ifelse(datafiltered2$`CRAFFT Score` <= 2, "Negative", "Positive")

craffttable1 <- table1fx(datafiltered2, quo(`CRAFFT Score`))

craffttable1

```

Alcohol & CRAFFT

```{r Alcohol & CRAFFT}

datafiltered2$alcCRAFFT = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`CRAFFT Score` == "Negative", "No Alc or No Subabuse", NA)

datafiltered2$alcCRAFFT = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`CRAFFT Score` == "Negative", "Yes Alc No Subabuse", datafiltered2$alcCRAFFT)

datafiltered2$alcCRAFFT = ifelse(datafiltered2$`Alcohol Use` == "No" & datafiltered2$`CRAFFT Score` == "Positive", "No Alc Yes Subabuse", datafiltered2$alcCRAFFT)

datafiltered2$alcCRAFFT = ifelse(datafiltered2$`Alcohol Use` == "Yes" & datafiltered2$`CRAFFT Score` == "Positive", "Yes Alc Yes Subabuse", datafiltered2$alcCRAFFT)

unique(datafiltered2$alcCRAFFT)

alcCRAFFTtable1 <- table1fx(datafiltered2, quo(alcCRAFFT))

alcCRAFFTtable1

```

Sexually Active
```{r Sexually Active}

sexactivtable1 <- table1fx(datafiltered, quo(`Sexually Active`))


sexactivtable1


```
Number of sex partners
```{r number of sex partners}

##Needs to be cleaned

sexnumtable1 <- table1fx(datafiltered, quo(`Number Of Sex Partner`))

sexnumtable1

```
Vaginal Sex
```{r Vaginal Sex}

vagsextable1 <- table1fx(datafiltered, quo(`Vaginal Sex`))

vagsextable1

```
Trusted Adult
```{r Trusted Adult}

trustadulttable1 <- table1fx(datafiltered, quo(`Trusted Adult`))

trustadulttable1

```
Contraception at last intercourse
```{r contraceptive at last intercourse}

datafiltered2$`No Contraceptive` = ifelse(datafiltered2$`No Contraceptive` == "uncheck", NA, datafiltered2$`No Contraceptive`)

contraceptivetable <- table1fx(datafiltered2, quo(`No Contraceptive`))


table(datafiltered2$`No Contraceptive`)

contraceptivetable

```
STD History
```{r STD History}

testcolumn = gsub('\\s+', '', datafiltered$`STD History`)

datafiltered2$`STD History` = case_when(
testcolumn=="chlamydiain9/14-txwithazithro"~"Yes",
testcolumn=="Yes(HSV1and2)"~"Yes",
testcolumn=="Yes(listtypeandtreatment)"~"Yes",
testcolumn=="Yes(listtypeandtreatment)--10/2015"~"Yes",
testcolumn=="Yes(listtypeandtreatment)-chlarmydia/zithromax3/2018"~"Yes",
testcolumn=="YesChlamydia"~"Yes",
testcolumn=="Yes,Chlamydia7/2015"~"Yes",
testcolumn=="No"~"No",
testcolumn=="no"~"No",
testcolumn=="No,gC/Chlamydianeg12/2017"~"No",
testcolumn=="no.negative7/2018"~"No")


STDtable1 <- table1fx(datafiltered2, quo(`STD History`))

STDtable1


```
univariate logistic regression

```{r univariate function}

datafiltered3 <- datafiltered2

datafiltered3$`PHQ9 Dx` =  ifelse(datafiltered3$`PHQ9 Dx` == "Positive", 1, 0)

table2fx <- function(data, col) {
  fit.covariate <- glm(data = data, `PHQ9 Dx` ~ col, family = "binomial")

  multireturn <- list(summary(fit.covariate), exp(fit.covariate$coefficients), exp(confint(fit.covariate)))
  return(multireturn)
}


```

```{r univariate logistic regression}

library(tidyverse)

fit.age <- table2fx(datafiltered3, (datafiltered3$AgeAsOf12312018))
fit.age

datafiltered3$BirthSex = as.factor(datafiltered3$BirthSex)
datafiltered3$BirthSex <- relevel(datafiltered3$BirthSex, ref = "M")
fit.sex <- table2fx(datafiltered3, as.factor(datafiltered3$BirthSex))
fit.sex

datafiltered3$`Smoking Status` = as.factor(datafiltered3$`Smoking Status`)
datafiltered3$`Smoking Status` <- relevel(datafiltered3$`Smoking Status`, ref = "Never Smoker")


fit.country <- table2fx(datafiltered3, as.factor(datafiltered3$`Birth Country`))
fit.country

fit.ageimmig <- table2fx(datafiltered3, (datafiltered3$`ageofimmigration`))
fit.ageimmig

fit.marijuana <- table2fx(datafiltered3, datafiltered3$`Marijuana Use`)
fit.marijuana

fit.alcohol <- table2fx(datafiltered3, datafiltered3$`Alcohol Use`)
fit.alcohol

fit.alcmarj <- table2fx(datafiltered3, datafiltered3$alcmarj)
fit.alcmarj

fit.druguse <- table2fx(datafiltered3, datafiltered3$`Drug Use`)
fit.druguse

fit.ecig <- table2fx(datafiltered3, datafiltered3$Ecigarette)
fit.ecig

datafiltered3$`Smoking Status` = as.factor(datafiltered3$`Smoking Status`)
datafiltered3$`Smoking Status` <- relevel(datafiltered3$`Smoking Status`, ref = "Never Smoker")
fit.smoke <- table2fx(datafiltered3, datafiltered3$`Smoking Status`)
fit.smoke

fit.craft <- table2fx(datafiltered3, as.factor(datafiltered3$`CRAFFT Score`))
fit.craft

fit.alcCRAFFT <- table2fx(datafiltered3, datafiltered3$`alcCRAFFT`)
fit.alcCRAFFT

fit.alcdrugs <-  table2fx(datafiltered3, datafiltered3$`alcdrugs`)
fit.alcdrugs

fit.sexactiv <- table2fx(datafiltered3, datafiltered3$`Sexually Active`)
fit.sexactiv

fit.trustadult <- table2fx(datafiltered3, datafiltered3$`Trusted Adult`)
fit.trustadult

fit.contracept <- table2fx(datafiltered3, datafiltered3$`No Contraceptive`)
fit.contracept

fit.std <- table2fx(datafiltered3, datafiltered3$`STD History`)
fit.std

fit.alcohol <- table2fx(datafiltered3, datafiltered3$`Alcohol Use`)
fit.alcohol 

```

multivariate regression

```{r}

fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + alcmarj + Ecigarette + `Smoking Status` + as.factor(`CRAFFT Score`) + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

exp(fit.multi$coefficients)

exp(confint(fit.multi))


```
Play with diff models
```{r}
##AIC 179.41 alcmarj 
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + alcmarj + Ecigarette + `Smoking Status` + `CRAFFT Score` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

exp(fit.multi$coefficients)

##AIC 173.91 alcmarj + drug use
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + alcmarj + `Drug Use` + Ecigarette + `Smoking Status` + as.factor(`CRAFFT Score`) + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

## AIC 184.09 alcCRAFFT
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Marijuana Use` + Ecigarette + `Smoking Status` + `alcCRAFFT` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

## AIC 176.56 alcCRAFFT + Drug use
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Marijuana Use` + `Drug Use` + Ecigarette + `Smoking Status` + `alcCRAFFT` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

##AIC 628.24 alcodruguse 
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + alcdrugs + `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

##AIC 176.71 alcodruguse + CRAFFT
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + alcdrugs + `Marijuana Use` + Ecigarette + `Smoking Status` + `CRAFFT Score` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

##AIC 175.05 alcohol + marijuana + drug use + crafft
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use` + `Drug Use` + `Marijuana Use` + Ecigarette + `Smoking Status` + `CRAFFT Score` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

##AIC 175.05 alcohol + marijuana + drug use + crafft
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Marijuana Use` + Ecigarette + `Smoking Status` + `CRAFFT Score` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

##AIC 1555.1 alcohol + marijuana + drug use  
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use` + `Drug Use` + `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active`, family = "binomial")
summary(fit.multi)
exp(fit.multi$coefficients)
exp(confint(fit.multi))

##AIC 1561.4 alcohol + marijuana 
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use`  + `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active`, family = "binomial")
summary(fit.multi)
exp(fit.multi$coefficients)
exp(confint(fit.multi))

##AIC 175.05 alcohol + marijuana + drug use  
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use` + `Drug Use` + `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active`, family = "binomial")
summary(fit.multi)
exp(fit.multi$coefficients)
exp(confint(fit.multi))

##AIC: 185.31 no alcohol, no drug use
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use`+ `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active` + `Trusted Adult`, family = "binomial")
summary(fit.multi)

exp(fit.multi$coefficients)

exp(confint(fit.multi))

fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use`+ `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active` + `CRAFFT Score` , family = "binomial")
summary(fit.multi)


```

```{r Final capstone model}


##AIC 1561.4 alcohol + marijuana 
fit.multi <- glm(data = datafiltered3, `PHQ9 Dx` ~ AgeAsOf12312018 + BirthSex + `Alcohol Use`  + `Marijuana Use` + Ecigarette + `Smoking Status` + `Sexually Active`, family = "binomial")
summary(fit.multi)
exp(fit.multi$coefficients)
exp(confint(fit.multi))

```

```{r indexing complete stuff}

# datafilteredcomplete = datafiltered3[complete.cases(datafiltered3 %>% select(AgeAsOf12312018, BirthSex, `Alcohol Use`, `Marijuana Use`, Ecigarette, `Smoking Status`, `Sexually Active`, `PHQ9 Dx`)),]
# 
# 
# dim(datafilteredcomplete)
```

